---
title: 'DATA 621: HW1'
output:
  pdf_document: default
---




# Data Exploration

## Correlations

We can start by plotting a correlation matrix, describing the pearson correlation coefficients between predictors. This will inform us about our dataset and provide insight when model building:

```{r, echo = F, warning = F, message = F}

library(tidyverse)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(knitr)

df.training = read_csv('moneyball-training-data.csv')

mat.correlation = cor(df.training, use = 'complete')
mat.correlation[upper.tri(mat.correlation)] = NA
mlt.correlation <- melt(mat.correlation)


ggplot(data = mlt.correlation, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 8, hjust = 1), axis.text.y = element_text(size = 8))+
 coord_fixed()
```

Both strikeouts by batters (`TEAM_BATTING_SO`) and errors (`TEAM_FIELDING_E`) are negatively correlated with the number of wins, while base hits by batters (`TEAM_BATTING_H`) is positively associated with the total win count.

We can further visualize the correlations against `TARGET_WINS` using scatterplots:

```{r, echo = F, warning = F, message = F}

mlt.training.plt = melt(select(df.training, -c("TARGET_WINS")), id.vars = c("INDEX"))
mlt.training.plt = merge(mlt.training.plt, select(df.training, c("INDEX", "TARGET_WINS")), by = "INDEX")
ggplot(data = mlt.training.plt, aes(value, TARGET_WINS)) + geom_point() + facet_wrap(~variable, scales = "free")
```

## Distributions

We can visualize the variables using histograms to account for non-normal distributions:

```{r, echo = F, warning = F, message = F}

mlt.training = melt(df.training, id.vars = "INDEX")
ggplot(data = mlt.training, aes(value)) + geom_histogram() + facet_wrap(~variable, scales = "free")
```

One of the key takeaways here is that strikeouts by batters has a bimodal distribution. Several variables, such as strikeouts by pitchers and walks allowed are skewed right. `TARGET_WINS` (i.e. our outcome of interest) has a normal distribution.



\newpage

# Appendix with R Code

```{r, eval = F, warning = F, message = F}

library(tidyverse)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(knitr)

df.training = read_csv('moneyball-training-data.csv')

# Correlation
mat.correlation = cor(df.training, use = 'complete')
mat.correlation[upper.tri(mat.correlation)] = NA
mlt.correlation <- melt(mat.correlation)


ggplot(data = mlt.correlation, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 8, hjust = 1), axis.text.y = element_text(size = 8))+
 coord_fixed()
```
